---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ChicagoHA

<!-- badges: start -->
<!-- badges: end -->

Explore and Import Chicago Health Atlas Data and Spatial Layers.

The Chicago Health Atlas is a collaboration between the Chicago Department of Public Health and PHAME Center at UIC, which aims to provide a place to "review, explore and compare health-related data over time and across communities." With the same spirit in mind, this tool aims to make Chicago health data analysis within R more painless and reproducable. For more information about the Chicago Health Atlas check out the [Chicago Health Atlas website](https://chicagohealthatlas.org/). 

**Note:** This is an unofficial R wrapper for the Chicago Health Atlas. This package is in no way affiliated with the Chicago Department of Public Health or PHAME Center at UIC.

## Installation

You can install the development version of ChicagoHA from
[GitHub](https://github.com/ryanzomorrodi/ChicagoHA) with:

``` r
# install.packages("pak")
pak::install("ryanzomorrodi/ChicagoHA")
```

## Usage
```{r}
library(ChicagoHA)
```

We can list all the topics (aka indicators) present within Chicago Health Atlas using `cha_topics()`. The most important column here is the `topic_key` which can be used to identify the topic within subsequent functions.
```{r}
cha_topics(progress = FALSE)
```

Then, we can explore what populations, time periods, and geographic scales that data is available for using `cha_coverage()`. Again, the most important columns here are the key columns which can be used to specify the data desired.
```{r}
cha_coverage("EDA", progress = FALSE)
```

Now, we can import our data using `cha_data()` specifying the keys we identified above.
```{r}
data <- cha_data(
  topic_key = "EDA",
  population_key = "",
  period_key = "2018-2022",
  layer_key = "neighborhood"
)
data
```

Let's create a map. But first, we will need to download the Community Areas geographic layer. We can do that with `cha_layer()`.
```{r}
layer <- cha_layer(
  layer_key = "neighborhood"
)
layer
```

Now we can make our map!
```{r message=FALSE}
library(dplyr)
library(ggplot2)

map_data <- layer |>
  left_join(data, join_by(geoid == neighborhood)) 

ggplot(map_data) +
  geom_sf(aes(fill = value), alpha = 0.7) +
  scale_fill_distiller(palette = "GnBu", direction = 1) +
  labs(
    title = "9th Grade Education Rate",
    fill = ""
  ) +
  theme_minimal()
```
